# Visualising and Exporting data{-}

```{r, include = FALSE}
source("common.R")
```

This section is intended for programmers and experts that would like to extend the capabilities of `sits`, either by including new data sources, ML algorithms, exporting data to be used in Python or QGIS, or including new display colors. 


## How colors work in sits{-}

In examples provided in the book, the color legend is taken for the predefined color table provided by `sits`.  This default color table is displayed using the command `sits_colors_show`. This color definition file assigns colors to 99 class names, including the IPCC and IGBP land use classes. 

```{r, tidy = "styler", out.width = "100%", out.height = "100%", echo = FALSE, fig.align="center", fig.cap="Default colors used in the sits package"}
# Display default sits colors
sits_colors_show()
```

The color table can be extended or adjusted by accessing and modifying the default color table, using the commands `sits_colors()` to retrieve the table and `sits_colors_set()` to update the table according to user choices, as shown in the example below.
```{r}
# retrieve the color table
color_tb <- sits_colors()
# show the color table
color_tb
```

As an example of a user-defined color table, consider a definition that covers level 1 of the Anderson Classification System used in the US National Land Cover Data, obtained by defining a new color table, as shown below. The colors can be defined by HEX values or by names accepted as R color codes.
```{r, tidy = "styler", out.width = "80%", out.height = "80%", fig.align="center", fig.cap="Example of Anderson Land Classification Scheme use in sits"}
# define a color table based on the Anderson Land Classification System
us_nlcd <- tibble::tibble(name = character(), color = character())
us_nlcd <- us_nlcd %>% 
  tibble::add_row(name = "Urban Built Up", color =  "#85929E") %>% 
  tibble::add_row(name = "Agricultural Land", color = "#F0B27A") %>% 
  tibble::add_row(name = "Rangeland", color = "#F1C40F") %>% 
  tibble::add_row(name = "Forest Land", color = "#27AE60") %>% 
  tibble::add_row(name = "Water", color = "#2980B9") %>% 
  tibble::add_row(name = "Wetland", color = "#D4E6F1") %>% 
  tibble::add_row(name = "Barren Land", color = "#FDEBD0") %>% 
  tibble::add_row(name = "Tundra", color = "#EBDEF0") %>% 
  tibble::add_row(name = "Snow and Ice", color = "#F7F9F9")
# load the color table into `sits`
sits_colors_set(us_nlcd)
# show the new color table used by sits
sits_colors_show()
```

```{r, eval = TRUE, echo = FALSE, warning = FALSE, message = FALSE} 
# reset the color table
sits_colors_reset()
```

As an alternative, users may define their own legends and pass them as parameters to to the `plot` function. Please see the example provided in Section "Map Reclassification" on Chapter "Image Classification in Data Cubes". 

## Exporting data to JSON{-}

Both the data cube and the time series tibble can be exported to exchange formats such as JSON.

```{r, tidy = "styler", eval = FALSE}
library(jsonlite)
# export the data cube to JSON
jsonlite::write_json(
  x = s2_20LKP_cube_MPC,
  path = "./data_cube.json",
  pretty = TRUE
)
# export the time series to JSON
jsonlite::write_json(
  x = samples_prodes_4classes,
  path = "./time_series.json",
  pretty = TRUE
)
```

